Reformulating phylogenetic mixed models to improve flexibility and speed

### 2019-07-08 phyloglmm_ms.tex
pushing again
puhsing yet again

In this paper, we will propose an alternative formulation of phylogenetic generalized linear mixed models that is mathematically equivalent to previous approaches, but more flexible.
Our new formulation can be incorporated in any framework that allows for random effects (such as random intercepts, slopes, and interactions) without the need to implement special correlation structures, by incorporating phylogenetic structures as part of the mean model. 
%% can be incorporated relatively easily into any mixed-model platform that allows new random-effects design matrices to be specified, and can be easily extended along with the features of mixed model such as multiple observational designs, phylogenetic random-slopes models, and multiple (nested or crossed) random effects.
We will compare our technique coded in R packages \pkg{lme4} and \pkg{glmmTMB} with existing R packages (i.e. \pkg{nlme} \citep{pinheiro2014r}, \pkg{phylolm} \citep{ho2014phylolm}, \pkg{pez} \citep{pearse2015pez} \pkg{phyr} (Li et al, unpublished), \pkg{MCMCglmm} \citep{hadfield2010mcmc}, and \pkg{brms} \citep{burkner2016brms}), fitting models to data from simulated model that incorporates phylogenetic variation from both predictors and tips/species as well as residual variation.
We expect that our approach will be fast as well as flexible, because \pkg{lme4} and \pkg{glmmTMB} use efficient approaches. 
%% TODO LAST LINE SEE PAGE 5

@@

We have presented the range of classical phylogenetic comparative methods (i.e. phylogenetic least squares, linear and mixed models) to fit models that incorporates different levels of phylogenetic variation.
Even within this scope, there is additional room for exploration we neglected, such as phylogenetic multivariate response models; non-Brownian evolutionary processes such as the Ornstein-Uhlenbeck (OU) model which accounts for both selection and drift processes \citep{butler2004phylogenetic}); Bayesian approaches \citep{hadfield2010general}; and variable-rate model, where evolutionary parameters vary across the phylogeny.
However, the simple approach we developed here offers an efficient way to handle phylogenetic comparative analysis for a wide range of univariate, Brownian-motion evolutionary models. 
This approach can in principle be combined flexibly with the state of the art phylogenetic mixed models using any platforms that supports independent latent variables such as Stan.
More importantly, this implementation in \pkg{lme4} and \pkg{glmmTMB} allows users to fit phylogenetic mixed models to the fullest (large data, unbalanced species observations, complex random effects) and explore new ideas.

>>>>

In this paper, we will propose an alternative formulation of phylogenetic generalized linear mixed models that is mathematically equivalent to previous approaches, but more flexible.
Our new formulation can be incorporated in any framework that allows for random effects (such as random intercepts, slopes, and interactions) without the need to implement special correlation structures, by incorporating phylogenetic structures as part of the mean model. 
%% can be incorporated relatively easily into any mixed-model platform that allows new random-effects design matrices to be specified, and can be easily extended along with the features of mixed model such as multiple observational designs, phylogenetic random-slopes models, and multiple (nested or crossed) random effects.
We will compare our technique coded in R packages \pkg{lme4} and \pkg{glmmTMB} with existing R packages (i.e. \pkg{nlme} \citep{pinheiro2014r}, \pkg{phylolm} \citep{ho2014phylolm}, \pkg{pez} \citep{pearse2015pez} \pkg{phyr} (Li et al, unpublished), \pkg{MCMCglmm} \citep{hadfield2010mcmc}, and \pkg{brms} \citep{burkner2016brms}), fitting models to data from simulated model that incorporates phylogenetic variation from both predictors and tips/species as well as residual variation.
% We expect that our approach will be fast as well as flexible, because \pkg{lme4} and \pkg{glmmTMB} use efficient approaches. 
%% TODO LAST LINE SEE PAGE 5

@@

We have presented the range of classical phylogenetic comparative methods (i.e. phylogenetic least squares, linear and mixed models) to fit models that incorporates different levels of phylogenetic variation.
Even within this scope, there is additional room for exploration we neglected, such as phylogenetic multivariate response models; non-Brownian evolutionary processes such as the Ornstein-Uhlenbeck (OU) model which accounts for both selection and drift processes \citep{butler2004phylogenetic}); Bayesian approaches \citep{hadfield2010general}; and variable-rate model, where evolutionary parameters vary across the phylogeny.
However, the simple approach we developed here offers an efficient way to handle phylogenetic comparative analysis for a wide range of univariate, Brownian-motion evolutionary models. 
This approach can in principle be combined flexibly with any platforms that supports independent latent variables such as Stan.
More importantly, this implementation in \pkg{lme4} and \pkg{glmmTMB} allows users to fit phylogenetic mixed models to the fullest (large data, unbalanced species observations, complex random effects) and explore new ideas.
